---
title: "DATA 609 HW Week 3 & 4"
author: "Ahsanul Choudhury"
date: "February 25, 2018"
output:
  html_document:
     theme: cerulean
     toc: true
     toc_float:
       collapsed: false
       smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 3

## Page 113: #2

**The following table gives the elongation $e$ in inches per inch (in./in.) for a given stress** ***S*** **on a streel wire measured in pounds per square inch ($lb/in^2$). Test the model $e=c_1S$ by potting the data. Estimate $c_1$ graphically.**

$S(\times 10^{-3})$|5|10|20|30|40|50|60|70|80|90|100
:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:
$e(\times 10^{5})$|0|19|57|94|134|173|216|256|297|343|390

***Solution***

To solve this problem we will create two models; one with median and one with buit in lm function in r.

```{r elongation, echo=TRUE}
if (!require('ggplot2')) install.packages('ggplot2')

S <- c(5, seq(10, 100, 10))
e <- c(0, 19, 57, 94, 134, 173, 216, 256, 297, 343, 390)


c <- e / S
df <- data.frame(S, e, c)

# Median model

c1 <- median(c)
c1

median_model <- ggplot(df) + geom_point(aes(x = S, y = e)) +  geom_abline(color = "red", slope = c1, intercept = 0) + ggtitle("Model with median c1")

median_model

# lm model forcing intercept = 0

lm <- lm(e ~ S + 0, df)
c1 <- lm$coefficients
c1


lm_model <- ggplot(lm) + geom_point(aes(x = S, y = e)) +  geom_abline(color = "red", slope = c1, intercept = 0) + ggtitle("Model with lm keeping intercept = 0")

lm_model

```

The lm model seems to fit better, so out estimated $c_1 \approx 3.7$ 


## Page 127: #10

**Data for planets**

```{r q_planets, echo=TRUE}
planets <- data.frame(Body = c("Mercury", "Venus", "Earth", "Mars", "Jupiter","Saturn", "Uranus", "Neptune"),
                      Period = c(7.6e6, 1.94e7, 3.16e7, 5.94e7, 3.74e8, 9.35e8, 2.64e9, 5.22e9),
                      Distance = c(5.79e10, 1.08e11, 1.5e11, 2.28e11, 7.79e11, 1.43e12,
                                   2.87e12, 4.5e12))
knitr::kable(planets)
```

**Fit the model** $y = ax^{3/2}$

***Solution***

We are looking for solution to least-squares formula $y = Ax^{n}$, where x = peroid and y = distance.

The formula in our case will be:

$$ a = \frac{\Sigma x_i^n y_i}{\Sigma x_i^{2n}} $$

Applying the given parameters we get:

$$a=\frac{\sum x_i^{\frac{3}{2}} y_i}{\sum x_i^{3}}$$


```{r planets, echo=TRUE}
a <- sum(planets$Period^(3/2)*planets$Distance)/sum(planets$Period^(3))
a

planets$y <- a * planets$Period^(3/2)

ggplot(planets, aes(x = Period, y = Distance)) + geom_point() + 
  geom_line(aes(x = Period, y = y), color = "red") + 
  labs(x = "Period (sec)", y = "Distance from Sun (m)")
```


***


# Week 4

## Page 191: #3

**Using Monte Carlo simulation, write an algorthm to calculate an approximation to $\pi$ by considering the number of random points selected inside the quarter circle**

$$ Q: x^2 + y^2 = 1, x \geq 0, y \geq 0 $$
**where the quarter circle is taken to be inside the square**

$$ S: 0 \leq x \leq i \; and \; 0 \leq y \leq 1$$

**Use the equation** $\pi/4 = area \; Q /area \; S$.

***Solution***

```{r monte_carlo, echo=TRUE}
# Function to estimate pi using sample size n
sim <- function(n) {
  x <- runif(n, 0, 1) # Random values for x between 0 and 1
  y <- runif(n, 0, 1) # Random values for y between 0 and 1
  Q <- x^2 + y^2 <= 1
  (sum(Q)/n) * 4
}

n <- c(10, 100, 1000, 10000)

set.seed(1023)

est_pi <- sapply(n, sim)

est_pi

```

## Page 194: #1

**Use the middle-square method to generate**

a.   **10 random numbers using $x_0$ = 1009.**

b.   **20 random numbers using $x_0$ = 653217.**

c.   **15 random numbers using $x_0$ = 3043.**

d.   **Comment about the results of each sequence. Was there cycling? Did each sequence degenerate rapidly?**

```{r}
if (!require('stringr')) install.packages('stringr')
set.seed(Sys.time())

# middle squre function

mid_square<- function(seed)
{
  length_seed <- nchar(seed) # find length of seed
  sq <- seed^2 # seed squre
  length <- nchar(sq)
  if(length < (length_seed * 2))
  {
    sq <- sprintf("%s%s", "0", sq) # add leading '0' where necessary
  }
  start <- (length_seed / 2) + 1 # start of middle
  end <- start + length_seed - 1 # end of middle
  mid_rnd <- str_sub(sq, start, end) # middle number
  return (as.numeric(mid_rnd))
}

# function to generate random number

rand <- function(n, seed)
{
  rand_num <- c()
  x0 <- seed
  for(i in 1:n)
  {
    x0 <- mid_square(x0)
    rand_num[i] <- x0
  }
  return (rand_num)
}
```

a.

```{r a, echo=TRUE}
rand(10, 1009)
```


b.


```{r b, echo=TRUE}
rand(20, 653217)
```


c.

```{r c, echo=TRUE}
rand(15, 3043)
```


d. Sequence "a"" degenarates rapidly within 5th iteration it gives a constant number. Sequence "b" has no degenarationor cycling issue and finally sequence "c" starts cycling after 7th iteration.