---
title: "DATA 609 HW Week 1 & 2"
author: "Ahsanul Choudhury"
date: "February 9, 2018"
output:
  html_document:
     theme: cerulean
     toc: true
     toc_float:
       collapsed: false
       smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1

## Page 8: #10

**Your grandparents have an annuity. The value of the annuity increases each month by an automatic deposit of 1% interest on the previous month's balance. Your grandparents withdraw $1000 at the beginning of each month for living expenses. Currently they have $50,000 in the annuity. Model the annuity with a dynamical system. Will the annuity run out of money? When?**

***Solution***

Let $a_n$ be the value of account after $n$ months

$a_{n+1}=1.01 \times a_n-1000$

$a_0 = 50000$

```{r annuity, echo=TRUE}
bal <- 50000
n <- 0
df <- data.frame(month = n, balance = bal)

# while loop until balance goes below $1000
while(bal > 1000){
  bal <- 1.01 * bal - 1000
  n <- n + 1
  df <- rbind(df, data.frame(month = n, balance = bal))
}
knitr::kable(df)
plot(df,type = "o", col = "red", main = "Grandparent's Annuity")
```

Annuity runs out in month `r (n)`

***

## Page 17: #9

**The data in the accompanying table show the speed $n$ (in increments of 5 mph) of an automobile and the associated distance $a_n$ in feet required to stop it once the brakes are applied. For instance, $n = 6$ (representing 6 × 5 = 30 mph) requires a stopping distances of $a_6 = 47 ft$.**

a.   **Calculate and plot the change $\Delta a_n$ versus nn. Does the graph reasonably approximate a linear relationship?**

b.   **Based on your conclusions in part (a), find a difference equation model for the stopping distance data. Test your model by plotting the errors in the predicted values against $n$. Discuss the appropriateness of the model.**

**n**|**1**|**2**|**3**|**4**|**5**|**6**|**7**|**8**|**9**|**10**|**11**|**12**|**13**|**14**|**15**|**16**
:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:
$a_n$|3|6|11|21|32|47|65|87|112|140|171|204|241|282|325|376


***Solution***

a.   We know,

$$\Delta a_n = a_{n+1}  - a_n$$

```{r data_table, echo=TRUE}
n <- (seq(1, 16))
an <- c(3, 6, 11, 21, 32, 47, 65, 87, 112, 140, 171, 204, 241, 282, 325, 376)
df <- data.frame(n, an)

# add column with delta_an
(df$delta_an <- c(diff(an), NA))

# plot
plot(df$n, df$delta_an, type = "o", xlab = "n", ylab = "delta_an")
```

Yes, the graph reasonably approximate a linear relationship.


b.   

```{r model, echo=TRUE}
# find intercept using lm fumction
df.lm <- lm(delta_an ~ n, data = df)
df.lm$coefficients
```

The difference equation model for the stopping distance data:

$$\Delta a_{n}= 3.246n - 1.105$$

or,

$$a_{n+1}= a_n + 3.246n - 1.105$$

```{r plot, echo=TRUE}
# functon for the model
model <- function(n, an, slope, intercept)
{
    est_an <- an + slope * n + intercept
    return(est_an)
}

# first observation will stay the same
n <- 1
est_an <- 3
my_est <- 3

# apply model function for remaining observations
for(i in 2:length(df$an))
{
  my_est[i] <- model(n,est_an, df.lm$coefficients[2], df.lm$coefficients[1])
  n <- n + 1
  est_an <- my_est[i]
}

df$est_an <-c(my_est)
df$est_an <- round(df$est_an, 3)
df$error <- c(df$an - df$est_an)
df
plot(df$n, df$error)

```
The model may not be appropriate, the errors do not appear to be normal, as the speed increases the error seems to increase aswell.

***

# Week 2

##  Page 69: #12

**A company with a fleet of trucks faces increasing maintenance costs as the age and mileage of the trucks increase**


*   Problem you would like to study: Benefit of replacing the fleet with a new on by leasing or financing vs operating with aging fleet. 

*   Variables:

      - Size of the fleet.
      
      - Operating cost of current fleet.
      
      - Salvage value of current fleet.
      
      - Financing cost for new fleet
       .
      - Operating cost of new fleet.
      
      - Depreciation
      
      - Savings on maintenance with new fleet.

*   Variable that may be neglected: For simplicity of the model difference in maintenance based on truck make and model may be neglected.

*   Variables might be considered as constants initially: Size of the fleet and fuel cost may be considered as constants initially.

*   Identify any submodels you would want to study in detail: Submodel I Would like to study is return on investment on each individual truck

*   Identify any data you would want collected: I would like to collect historical data on maintenance with age and mileage.


***

## Page 79: #11

$y alpha x^3$

y|0|1|2|6|14|24|37|58|82|114
:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:
x|1|2|3|4|5|6|7|8|9|10

***Solution***

```{r q11, echo=TRUE}
y <- c(0,1,2,6,14,24,37,58,82,114)
x <- c(seq(1,10))
dfq11 <- data.frame(x,y)

dfq11$x3 <- x^3

dfq11$k <- dfq11$y/dfq11$x3

dfq11$prop <- dfq11$x3 * dfq11$k

dfq11$m_k <- mean(dfq11$k)
pred <-  dfq11$m_k * dfq11$x3

library(ggplot2)
ggplot(dfq11) + geom_line(aes(x, y), colour="blue") + geom_line(aes(x, y=pred), colour="red") + labs(title="Proportionality, Blue = Actual, Red = Predicted")
```

This data set supports the stated proportionality model.