---
title: "Ahsanul_Choudhury_DATA_609_Project"
author: "Ahsanul Choudhury"
date: "May 15, 2018"
output:
  pdf_document: default
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage[normalem]{ulem}
---
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\tableofcontents
\newpage

```{r load_libraries, include=FALSE, cache=FALSE}
if (!require('readxl')) install.packages('readxl')
if (!require('httr')) install.packages('httr')
if (!require('dplyr')) install.packages('dplyr')
if (!require('reshape2')) install.packages('reshape2')
if (!require('RCurl')) install.packages('RCurl')
if (!require('ggplot2')) (install.packages('ggplot2'))
if (!require('gridExtra')) (install.packages('gridExtra'))

```

# Introduction

According to a report titled "The number of hate crimes rose in 2016" published on CNN.com (https://www.cnn.com/2017/11/13/politics/hate-crimes-fbi-2016-rise/index.html) on November 13, 2017, hate crimes in the United States have increased to a point not seen in recent history. They showed some statistics from Federal Bureau of Investigation and used graphics to back their claim. On a graph the visualized how hate crime spiked around the time of November 2016 election. One of the reason put forward for this is share of proverty in the majority race member of the population and which was blamed on influx of immigrant by some political leaders during campaigning.  

My Goal for this project is to collect data form various sources and construct a mathematical model to see the relationship between the share of poverty in white population and reported hate crimes in each state.


# Data Collection and Preparation

To collect my data I have used the following three sources:

https://ucr.fbi.gov/hate-crime/2016/tables/table-11

https://www.census.gov/data/datasets/2016/demo/popest/state-total.html

https://github.com/fivethirtyeight/data/blob/master/hate-crimes/hate_crimes.csv

Data from these source downloaded cleaned and combined to create my final dataset. All the information for the state of Hawaii was not available so it was left out for this project. The following table shows the final dataset; the first column represent the states, second column represents the Share of white residents who are living in poverty and the third column represents the number of hate crimes reported by FBI for the year of 2016. I chose to convert the reported hate crime for every 100,000 of population due large variance in total population in each state. 


```{r}
#Download and read data
url <- "https://github.com/choudhury1023/DATA-609/raw/master/Project/table_11_offenses_offense_type_by_participating_state_2016.xls"
GET(url, write_disk("2016_data.xls", overwrite=TRUE))
df <- read_excel("2016_data.xls")

#Data Cleanup and manipulation

df <- df[-c(1:6, 57:60), ]

df1 <- df[ ,-c(3:17)]

colnames(df1) <- c("state","hate.crimes")

write.csv(df1, "df1.csv")

```




```{r}
url <- c("https://github.com/choudhury1023/DATA-609/raw/master/Project/nst-est2016-01.xlsx")
GET(url, write_disk("population.xlsx", overwrite=TRUE))
df6 <- read_excel("population.xlsx")

df6 <- df6[-c(1:8, 20, 60:67), ]
df6 <- df6[-c(2:9)]
colnames(df6) <- c("state","population")

write.csv(df6, "pop.csv")
```



```{r}
raw <- getURL("https://raw.githubusercontent.com/choudhury1023/DATA-609/master/Project/hate_crimes.csv")
df7 <- read.csv(text = raw)
df7 <- df7[-c(2:6, 8:12)]
df7 <- df7[-c(12), ]


df8 <- bind_cols(df1, df6, df7)
df8 <- transform(df8, hate.crimes = as.numeric(hate.crimes))
df8$avg_hatecrimes_per_100k_fbi_2016 <- (df8$hate.crimes/df8$population)*100000
df8 <- transform(df8, share_white_poverty = as.numeric(share_white_poverty))
df8 <- transform(df8, avg_hatecrimes_per_100k_fbi_2016 = as.numeric(avg_hatecrimes_per_100k_fbi_2016))
final_df <- df8[-c(2:5)]
final_df1 <- df8[-c(2:5)]
final_df2 <- df8[-c(2:5)]

knitr::kable(final_df)
```

# Model Construction


## Geometric Similarity Model

Assumptions:

We assume that the share of white poverty is proportional to any characteristic dimension cubed.

$$V \alpha l^3$$

If we assume a constant share of white poverty, then the share state's number of hate crimes reported  is proportional to it's share of white poverty,

$$V \alpha W$$

$$V \alpha W \alpha l^3$$

Let the characteristic dimension of l be the number of hate crime reported, which is chosen because it is reported as an indicator of share of white poverty.

$$W = kl^3, k > 0$$


```{r}

final_df$z1 = c(NA,tail(final_df$share_white_poverty,-1)-head(final_df$share_white_poverty,-1))
final_df$z2= c(NA,tail(final_df$avg_hatecrimes_per_100k_fbi_2016,-1)-head(final_df$avg_hatecrimes_per_100k_fbi_2016,-1))

# finding the slope
final_df$slope<- final_df$z1 / final_df$z2

slope<- mean(final_df$slope, na.rm= TRUE)
slope

# calculating f(x) using the slope 
final_df$fx<- slope*(final_df$avg_hatecrimes_per_100k_fbi_2016)^(-1/3)


p1<- ggplot(data = df, aes(x = final_df$avg_hatecrimes_per_100k_fbi_2016, y = final_df$share_white_poverty)) + 
  labs(x = "Hate Crimes 
       Reported per 100k", y = "Share of white residents 
       who are living in poverty") +
  geom_line(color = "red")+
  ggtitle("Hate Crimes Reported 
per 100k Share of white
poverty (Original data)")

p2<- ggplot(data = df, aes(x = final_df$avg_hatecrimes_per_100k_fbi_2016, y = final_df$fx)) + 
  labs(x = "Hate Crimes 
       Reported per 100k", y = "Share of white residents 
       who are living in poverty") +
  geom_line(color = "red")+
  ggtitle("Hate Crimes Reported 
per 100k Share of white
poverty (Geometric Similarity 
Fitted model)")

grid.arrange(p1, p2, nrow=2, ncol=1)

```


## Least-Squares Criterion Model

Our formula for Least-Squares Criterion Model:

$$a = \frac{\Sigma x_i^n y_i}{\Sigma x_i^{2n}}$$
Where $x$ = avg_hatecrimes_per_100k_fbi_2016 and $y$ = share_white_poverty

$n$ is fixed to given collection of data points


```{r}
a <- sum(final_df1$avg_hatecrimes_per_100k_fbi_2016^(-1/3)*final_df1$share_white_poverty)/sum(final_df1$avg_hatecrimes_per_100k_fbi_2016^(-2/3))
a

final_df1$share_white_poverty_pred <- a * final_df1$avg_hatecrimes_per_100k_fbi_2016^(-1/3)


ggplot(final_df1, aes(x = avg_hatecrimes_per_100k_fbi_2016, y = share_white_poverty)) + geom_point() + 
  geom_line(aes(x = avg_hatecrimes_per_100k_fbi_2016, y = share_white_poverty_pred), color = "red") + 
  labs(x = "Hate Crimes Reported per 100k", y = "Share of white residents who are living in poverty")

```




# Conclusion

```{r}
p3<- ggplot(data = final_df1, aes(x = final_df1$avg_hatecrimes_per_100k_fbi_2016, y = final_df1$share_white_poverty_pred)) + 
  labs(x = "Hate Crimes 
       Reported per 100k", y = "Share of white residents 
       who are living in poverty") +
  geom_line(color = "red")+
  ggtitle("Hate Crimes Reported 
per 100k Share of white poverty
(Least-Squares Criterion Fitted model)")

grid.arrange(p1, p2, p3, nrow=2, ncol=2)
```

The models do not fit tightly with our observed data. Hate crimes alone can not be a predictor of poverty in a portion of a state's population and poverty by itself can not be the sole contributing factor to hate crime. Also, our data have some outliers, for example District of Columbia has a large number of hate crimes reported in 2016 and has a relatively low share of poverty compared to other states. There are other socioeconomic factors which may be a better indicator of poverty in a segment of population like level of education or combination of various other factor. We started the project to see the relationship between hate crimes reported in a state and share of poverty in a segment of population and we are concluding this project by saying the relationship is complex and not truly linear.


# Reference

Giordano, F. R., Fox, W. P., & Horton, S. B. (2014). A first course in mathematical modeling. Australia: Brooks/Cole, Cengage Learning.

Table 11. (2017, November 03). Retrieved from https://ucr.fbi.gov/hate-crime/2016/tables/table-11

https://www.census.gov/data/datasets/2016/demo/popest/state-total.html

https://github.com/fivethirtyeight/data/blob/master/hate-crimes/hate_crimes.csv



